Drought vulnerability of tree growth

### 2024-07-02 scripts/explore_analysis.Rmd
updated figures and explore analysis

pre_yr_census<-ggplot()+
  geom_area(data=chirps_hkk, aes(x=time, y=precipitation), fill="blue", alpha=0.5)+
  #geom_line(data=climate_data_orig[climate_data_orig$variable=="tmx",], aes(x=as.Date(paste0(year, "-", month, "-", "01"), format = "%Y-%m-%d"), y=value*5), col="red")+
  geom_rect(data=census_int, aes(xmin = start, xmax = end, ymin = -Inf, ymax = Inf), 
              fill = "grey10", alpha = 0.5) +
  geom_text(data=census_int, aes(x=start, y=500, 
                                 label= paste0("C", 1:29), angle=90), size=2)+
  # scale_y_continuous(    # Features of the first axis
  #   name = "monthly rainfall (mm)",
  #   
  #   # Add a second axis and specify its features
  #   sec.axis = sec_axis(~.*0.2, name="maximum temperature (Celsius °)"))+
  xlab("")+theme_bw()+ ylab("total monthly precipitation (mm)")+
  ggtitle("HKK dendroband censuses and CHIRPS rainfall data")+
  #scale_x_continuous(n.breaks=15)+
  theme(axis.text.x = element_text(angle=45))

@@

dendro_2010<- dendro.data.noouts %>%
  filter(Cno %in% c(1, 3, 5))%>%
  group_by(Tag) %>% #grouping by treeID to calculate increment
  mutate(calcDBH_min1=dplyr::lag(calcDBH, n=1, default=NA, order_by=Tag),
         cii_min1=dplyr::lag(Cii, n=1, default=NA, order_by=Tag),
         tdif=difftime(Date2, lag(Date2, n=1, default=NA, order_by=Tag), units = "days"),
         inc=calcDBH-calcDBH_min1,
         inc_annual=(inc/as.numeric(tdif))*365,
         inc.time=year(Date2)-lag(year(Date2)),
         #removing increment value if the previous measure was before the last year
         #inc=ifelse(inc.time==1, inc, NA),
         inc_annual=ifelse(inc.time==1, inc_annual, NA),
         treeID=paste0("HKK.", Tag))%>%
  filter(!is.na(inc))%>%
  filter(!is.na(inc_annual))%>%
  ungroup()%>%
  dplyr::mutate(
    diainc.scaled = scale(inc_annual),
    out.inc = ifelse(abs(diainc.scaled)>3, 1, 0))#%>%
  #filter(out.inc==0)

@@

dendro_2015<- dendro.data.noouts %>%
  filter(Cno %in% c(11, 13, 15))%>%
  group_by(Tag) %>% #grouping by treeID to calculate increment
  mutate(calcDBH_min1=dplyr::lag(calcDBH, n=1, default=NA, order_by=Tag),
         cii_min1=dplyr::lag(Cii, n=1, default=NA, order_by=Tag),
         tdif=difftime(Date2, lag(Date2, n=1, default=NA, order_by=Tag), units = "days"),
         inc=calcDBH-calcDBH_min1,
         inc_annual=(inc/as.numeric(tdif))*365,
         inc.time=year(Date2)-lag(year(Date2)),
         #removing increment value if the previous measure was before the last year
         #inc=ifelse(inc.time==1, inc, NA),
         inc_annual=ifelse(inc.time==1, inc_annual, NA),
         treeID=paste0("HKK.", Tag))%>%
  filter(!is.na(inc))%>%
  filter(!is.na(inc_annual))%>%
  ungroup()%>%
  dplyr::mutate(
    diainc.scaled = scale(inc_annual),
    out.inc = ifelse(abs(diainc.scaled)>3, 1, 0))%>%
  filter(out.inc==0)

@@

library(gridExtra)
png("../doc/display/increments_raw.png", width=8, height=4, units="in", res=300)
grid.arrange(dendro_2010_plot, dendro_2015_plot, nrow=1)
dev.off()

```
```{r drought resistance}

@@

resistance_2010 <- dendro_2010 %>%
  group_by(Tag) %>% #grouping by treeID to calculate increment
  mutate(inc_min1=dplyr::lag(inc_annual, n=1, default=NA, order_by=Tag),
         resistance.div = inc_annual/inc_min1,
         resistance.dif = inc_annual-inc_min1)%>%
  filter(!is.na(inc_min1), !is.infinite(resistance.div))

resistance_2010<-merge(resistance_2010, dec_williams, by="sp", all.x=T)

@@

resistance_2015 <- dendro_2015 %>%
  group_by(Tag) %>% #grouping by treeID to calculate increment
  mutate(inc_min1=dplyr::lag(inc_annual, n=1, default=NA, order_by=Tag),
         resistance.div = inc_annual/inc_min1,
         resistance.dif = inc_annual-inc_min1)%>%
  filter(!is.na(inc_min1), !is.infinite(resistance.div))

resistance_2015<-merge(resistance_2015, dec_williams, by="sp", all.x=T)

@@

png("../doc/display/resistance_dif.png", width=8, height=4, units="in", res=300)
grid.arrange(resistance_2010_dif, resistance_2015_dif, nrow=1)
dev.off()

```
```{r basic model}

@@

resistance_all<-rbind(resistance_2010, resistance_2015)
resistance_all$yr<-c(rep(2010, nrow(resistance_2010)), rep(2015, nrow(resistance_2015)))

resistance_model<-resistance_all %>%
  select(resistance.dif, calcDBH_min1, cii_min1, williams_dec, yr)

@@

library(nlme)
library(ggplot2)

@@

```{r maps of species with tree ring records, echo=F, eval=F}
library(ggplot2)

pat<-as.character(read.table("git_pat.txt")[1,1])

remotes:::download("stem1.RData",
                   "https://raw.githubusercontent.com/forestgeo/HKK-tree-growth-data/main/data/Census/raw_data/hkk.stem1.rdata",
                   auth_token = pat)
remotes:::download("stem2.RData",
                   "https://raw.githubusercontent.com/forestgeo/HKK-tree-growth-data/main/data/Census/raw_data/hkk.stem2.rdata",
                   auth_token = pat)
remotes:::download("stem3.RData",
                   "https://raw.githubusercontent.com/forestgeo/HKK-tree-growth-data/main/data/Census/raw_data/hkk.stem3.rdata",
                   auth_token = pat)
remotes:::download("stem4.RData",
                   "https://raw.githubusercontent.com/forestgeo/HKK-tree-growth-data/main/data/Census/raw_data/hkk.stem4.rdata",
                   auth_token = pat)

load("stem1.RData")
load("stem2.RData")
load("stem3.RData")
load("stem4.RData")

hkk_census <- hkk_census %>%
  filter(status=="A")

hkk_census_ring_sp<-hkk_census %>%
  filter(CensusID==3)%>%
  filter(sp %in% ring.sp)

hkk_census_sp <- hkk_census %>%
  filter(CensusID==4)%>%
  #group_by(sp, tag, CensusID) %>%
  #dplyr::summarise(tot.agb=sum(agb, na.rm=T)) %>%
  group_by(sp, tag)%>%
  dplyr::summarise(tot.agb=sum(agb, na.rm=T)) %>%
  group_by(sp) %>%
  dplyr::summarise(n.tree=length(unique(tag)),
                   sum.agb=sum(tot.agb))%>%
  ungroup()%>%
  dplyr::mutate(pct.agb=sum.agb*100/sum(sum.agb, na.rm=T))%>%
  arrange(-pct.agb)%>%
  dplyr::mutate(pct.agb.cum=cumsum(pct.agb))

sp_top15<-c("SACCLI", "HOPEOD", "POLYVI", "TETRNU", "VATIHA",
            "ALPHVE", "DIPTAL", "DIOSWI", "ARYTLI", "BACCRA",
            "GARCSP", "CROTRO", "GARUPI", "NEOLOB", "PHOEPA")

#pct agb of the top 15 dendro species
sum(hkk_census_sp$pct.agb[hkk_census_sp$sp %in% sp_top15])

## Fastest and slowest in 2015 and 2016

dendro.data.noouts <- dendro.data.noouts %>%
  group_by(year) %>% 
  dplyr::mutate(growth_quantile = ntile(diainc, 4),
                resid_quantile = ntile (inc.resid, 4))

growth_2015 <- dendro.data.noouts %>%
  filter(year==2015)

quantiles_2015<-quantile(growth_2015$diainc, probs=c(0.25, 0.5, 0.75, 1))

growth_2015_plot<-ggplot(growth_2015, aes(x=diainc)) +
  geom_density(fill="brown", col="brown", alpha=0.5) +
  geom_vline(xintercept=quantiles_2015, col="brown")+ theme_bw()

growth_2014 <- dendro.data.noouts %>%
  filter(Cno==11)

quantiles_2014<-quantile(growth_2014$diainc, probs=c(0.25, 0.5, 0.75, 1))

growth_2016 <- dendro.data.noouts %>%
  filter(year==2016)

quantiles_2016<-quantile(growth_2016$diainc, probs=c(0.25, 0.5, 0.75, 1))

growth_2016_plot<-ggplot(growth_2016, aes(x=diainc)) +
  geom_density(fill="brown", col="brown", alpha=0.5) +
  geom_vline(xintercept=quantiles_2016, col="brown")+ theme_bw()

growth_2017 <- dendro.data.noouts %>%
  filter(year==2017)

quantiles_2017<-quantile(growth_2017$diainc, probs=c(0.25, 0.5, 0.75, 1))

growth_2017_plot<-ggplot(growth_2017, aes(x=diainc)) +
  geom_density(fill="brown", col="brown", alpha=0.5) +
  geom_vline(xintercept=quantiles_2016, col="brown")+ theme_bw()

growth_2014_2015_plot<-ggplot() +
  geom_density(data = growth_2014, aes(x=diainc), fill="forestgreen", col="forestgreen", alpha=0.5) +
  geom_vline(xintercept=quantiles_2014, col="forestgreen")+
  geom_density(data= growth_2015, aes (x=diainc), fill="brown", col="brown", alpha=0.5) +
  geom_text(aes(x=0.75, y=1, label="2014"), col="forestgreen", size=3)+
  geom_text(aes(x=0.75, y=1.5, label="2015"), col="brown", size=3)+
  geom_vline(xintercept=quantiles_2015, col="brown")+ 
  xlab("diameter increment (cm)") + ylab("density")+
  theme_bw()

dendro.data.noouts$year_fac<-as.character(dendro.data.noouts$year)
dendro.data.noouts$year_fac<-factor(dendro.data.noouts$year_fac, levels=as.character(rev(2009:2023)))

growth_all_years<-ggplot(dendro.data.noouts, aes(x = diainc, y = year_fac, 
                       fill = stat(x))) +
  geom_density_ridges_gradient(scale = 3, rel_min_height = 0.01) +
  scale_fill_viridis(name = "growth", option = "H", direction=-1) +
  labs(title = 'tree growth in HKK (cm)', y="year") +
  #theme_ipsum() +
  geom_vline(xintercept=0, lty=2)+
  theme_bw()+
    theme(
      legend.position="none",
      panel.spacing = unit(0.1, "lines"),
      strip.text.x = element_text(size = 8)
    )

growth_all_years

png("../doc/display/growth_2014_2015.png", width=4, height=4, units="in", res=300)
growth_2014_2015_plot
dev.off()

png("../doc/display/growth_allyrs.png", width=4, height=4, units="in", res=300)
growth_all_years
dev.off()

@@

## Size classes

size_top10<-ggplot(dendro.top10.dat, aes(x=dbh))+
  geom_density()+
  geom_vline(aes(xintercept=35))+
  geom_vline(aes(xintercept=75))

size_top10

#trying three categories
# <50, 50-100, >100

dendro.top10.dat<-dendro.top10.dat %>%
  dplyr::mutate(size=ifelse(dbh<35, "< 35 cm dbh",
                       ifelse(dbh<70, "35-75 cm dbh", ">75 cm dbh")))

growth_sp10<-ggplot(dendro.top10.dat, aes(x=factor(sp, levels=sp_top10), y=diainc,
                                              #color=factor(sp),
                                          #group=factor(size, levels=c("small", "medium", "large")),
                                              fill=factor(sp, levels=sp_top10)))+ 
  #geom_violin(position = "dodge")+ 
  geom_boxplot(position = "dodge")+ 
  ylab("Annualised diameter increment (cm)")+theme_bw()+
  #geom_hline(yintercept = mean(dendro.top10.dat$diainc, na.rm=T))+
  #geom_hline(yintercept = median(dendro.top10.dat$diainc, na.rm=T), lty=2)+
  geom_hline(yintercept = 0, lty=3)+
  facet_grid(factor(size, levels=c("< 35 cm dbh", "35-75 cm dbh", ">75 cm dbh"))~.)+
      theme(axis.text.y = element_blank())+
  scale_fill_viridis_d()+
  scale_color_viridis_d()+
  ggtitle("HKK dendroband data from 2008 to 2023")+
  guides(fill=guide_legend(title="species"),
         color=guide_legend(title="species"))+coord_flip()

growth_sp10

growth_sp10<-ggplot(dendro.top10.dat, aes(x=factor(size, levels=c("< 35 cm dbh", "35-75 cm dbh", ">75 cm dbh")), y=diainc,
                                              #color=factor(sp),
                                          #group=factor(size, levels=c("small", "medium", "large")),
                                              fill=factor(sp, levels=sp_top10)))+

  geom_violin(position = "dodge")+ 
  geom_boxplot(position = "dodge", col="black", fill="white", width=0.1)+ 
  ylab("Annualised diameter increment (cm)")+theme_bw()+
  #geom_hline(yintercept = mean(dendro.top10.dat$diainc, na.rm=T))+
  #geom_hline(yintercept = median(dendro.top10.dat$diainc, na.rm=T), lty=2)+
  geom_hline(yintercept = 0, lty=3)+
  #facet_grid(factor(size, levels=c("< 35 cm dbh", "35-75 cm dbh", ">75 cm dbh"))~.)+
      theme(axis.text.x = element_text(angle=45))+
  xlab("")+
  facet_grid(.~factor(sp, levels=sp_top10))+
  scale_fill_viridis_d()+
  scale_color_viridis_d()+
  ggtitle("HKK dendroband data from 2008 to 2023")+
  guides(fill=guide_legend(title="species"),
         color=guide_legend(title="species"))#+coord_flip()

growth_sp10

## Canopy illumination

cii_top10<-ggplot(dendro.top10.dat, aes(x=cii))+
  geom_histogram()

cii_top10

#trying three categories
# <50, 50-100, >100

#dendro.top10.dat<-dendro.top10.dat %>%
#  dplyr::mutate(size=ifelse(dbh<50, "small",
#                       ifelse(dbh<100, "medium", "large")))

growth_sp10_cii<-ggplot(dendro.top10.dat[dendro.top10.dat$cii %in% (1:5),], 
                        aes(x=factor(sp, levels=sp_top10), y=diainc,
                                              #color=factor(sp),
                                          #group=factor(size, levels=c("small", "medium", "large")),
                                              fill=factor(sp, levels=sp_top10)))+ 
  #geom_violin(position = "dodge")+ 
  geom_boxplot(position = "dodge")+ 
  ylab("Annualised diameter increment (cm)")+theme_bw()+
  #geom_hline(yintercept = mean(dendro.top10.dat$diainc, na.rm=T))+
  #geom_hline(yintercept = median(dendro.top10.dat$diainc, na.rm=T), lty=2)+
  geom_hline(yintercept = 0, lty=3)+
  facet_grid(cii~.)+
      theme(axis.text.y = element_blank())+
  ggtitle("HKK dendroband data from 2008 to 2023 - canopy illumination")+
  scale_fill_viridis_d()+
  scale_color_viridis_d()+
  guides(fill=guide_legend(title="species"),
         color=guide_legend(title="species"))+coord_flip()

growth_sp10_cii

growth_sp10_cii<-ggplot(dendro.top10.dat[dendro.top10.dat$cii %in% (1:5),], 
                        aes(
                          #x=factor(sp, levels=sp_top10), 
                          x=factor(cii), 
                            y=diainc,
                                              #color=factor(sp),
                                          #group=factor(size, levels=c("small", "medium", "large")),
                                              fill=factor(sp, levels=sp_top10)))+ 
  #geom_violin(position = "dodge")+ 
  geom_boxplot(position = "dodge")+ 
  ylab("Annualised diameter increment (cm)")+theme_bw()+
  #geom_hline(yintercept = mean(dendro.top10.dat$diainc, na.rm=T))+
  #geom_hline(yintercept = median(dendro.top10.dat$diainc, na.rm=T), lty=2)+
  geom_hline(yintercept = 0, lty=3)+
  facet_grid(.~factor(sp, levels = sp_top10))+
      theme(axis.text.y = element_blank())+
  ggtitle("HKK dendroband data from 2008 to 2023 - canopy illumination")+
  scale_fill_viridis_d()+
  scale_color_viridis_d()+
  #coord_flip()+
  guides(fill=guide_legend(title="species"),
         color=guide_legend(title="species"))
growth_sp10_cii

```
## spatial patterns

```{r spatial patterns, echo=F}

trees.top10 <- trees.top10 %>%
  dplyr::mutate(treeID=paste0("HKK.", Tag),
                plotID=paste0(Col, "_", Row))

dendro.top10.dat<-merge(dendro.top10.dat, 
                        dplyr::select(trees.top10, c("Col", "Row", "treeID", "X", "Y", "plotID")), 
                        by="treeID", all.x=T)

dendro.top10.dat<-dendro.top10.dat %>%
  dplyr::mutate(plotID=paste0(Col, "_", Row))

census_spatial<-hkk_census %>%
  group_by(quadrat) %>%
  dplyr::summarise(sp.no=length(unique(sp)),
                   sp.top.10 = length(unique(sp %in% sp_top10)))%>%
  dplyr::mutate(Col=as.numeric(str_sub(quadrat, 1, 2)),
                Row=as.numeric(str_sub(quadrat, 3, 4)))

tiles<-ggplot(census_spatial, aes(x=Col, y=Row, fill=sp.no))+
  geom_tile()+
  #facet_wrap(~year)+
  coord_fixed()+
  #xlim(c(0,49))+ylim(c(0, 24))+
  theme_bw()

tiles

tiles<-ggplot(census_spatial, aes(x=Col, y=Row, fill=sp.top.10))+
  geom_tile()+
  #facet_wrap(~year)+
  coord_fixed()+
  #xlim(c(0,49))+ylim(c(0, 24))+
  theme_bw()

tiles

spatial_summary<- dendro.top10.dat %>%
  group_by(Col, Row, plotID, sp, year) %>%
  dplyr::summarise(inc.sum=sum(inc.resid, na.rm=T)) %>%
  group_by(Col, Row, plotID, year) %>%
  dplyr::summarise(inc.diff=var(inc.sum, na.rm=T))

tiles<-ggplot(spatial_summary, aes(x=Col, y=Row, fill=log(inc.diff, 10)))+
  geom_tile()+
  facet_wrap(~year)+coord_fixed()+
  xlim(c(0,49))+ylim(c(0, 24))+
  theme_bw()

tiles

## evergreen/dec

```{r evg/dec, echo=F}

sp_top15<-c("SACCLI", "HOPEOD", "POLYVI", "TETRNU", "VATIHA", 
            "ALPHVE", "DIPTAL", "DIOSWI", "ARYTLI", "BACCRA",
            "GARCSP", "CROTRO", "GARUPI", "NEOLOB", "PHOEPA")

#these 15 species had the most dendrobands and together made up 51.22% of the agb in 2009

#below values eyeballed from Williams et al. 2008 Fig 1.
sp_top15_phenoscore<-c(2.5, 0.8, 1.3, 4, 0.7, 
                       1, 1.25, 1.27, 0.2, 0.3,
                       0.4, 0.65, 3.95, 0.15, 0.25)

#no info on Alphonsea ventricosa in the paper. 
#Apparently evergreen, so used 1 as value - middle of evergreen range

pheno<-as.data.frame(cbind(sp_top15, sp_top15_phenoscore))
colnames(pheno)<-c("sp", "phenoscore")

dendro.top10.dat<-merge(dendro.top10.dat, pheno, by="sp")

#categorising as evg and deciduous
dendro.top10.dat$cat<-ifelse(dendro.top10.dat$phenoscore<=1, "evg", "dec")

growth_yr_cat_detrend<-ggplot(dendro_top15, aes(x=factor(cat), y=inc.resid,
                                              #group=c(year, sp),
                                              #group=sp,
                                              fill=factor(cat)))+ 
  geom_violin(position = "dodge")+ 
  ylab("Residuals of annualised diameter increment")+theme_bw()+
  facet_grid(.~year)+
  geom_hline(yintercept = 0)+
  scale_fill_manual(values=c("sienna4", "darkgreen"))+
  theme(axis.text.x = element_blank())

growth_yr_cat_detrend

>>>>

spei_hkk<-read.csv("../data/climate/SPEI_HKK_from_GEE.csv")
colnames(spei_hkk)[1]<-"time"

#making dates manageable
spei_hkk<- spei_hkk %>%
  dplyr::mutate(time=as.Date(time, format = "%b %d, %Y"),
                year=year(time),
                month=month(time))

pre_yr_census<-ggplot()+
  #geom_area(data=chirps_hkk, aes(x=time, y=precipitation), fill="blue", alpha=0.5)+
  geom_line(data=chirps_hkk, aes(x=time, y=precipitation), col="blue", alpha=0.5)+
  geom_line(data=spei_hkk, aes(x=time, y=(100*SPEI_01_month-250)), col="red")+
  geom_rect(data=census_int, aes(xmin = start, xmax = end, ymin = -Inf, ymax = Inf), 
              fill = "grey10", alpha = 0.5) +
  geom_text(data=census_int, aes(x=start, y=500, 
                                 label= paste0("C", 1:29), angle=90), size=2)+
   scale_y_continuous(    # Features of the first axis
     name = "monthly rainfall (mm)",

     # Add a second axis and specify its features
     sec.axis = sec_axis(~(.+250)*0.01, name="SPEI"))+
  xlab("")+theme_bw()+ ylab("total monthly precipitation (mm) from CHIRPS")+
  ggtitle("HKK dendroband censuses and climate")+
  #scale_x_continuous(n.breaks=15)+
  theme(axis.text.x = element_text(angle=45))

@@

dendro_2010<- dendro.data.noouts %>%
  filter(Cno %in% c(1, 3, 5))%>%
  group_by(Tag) %>% #grouping by treeID to calculate increment
  mutate(calcDBH_min1=dplyr::lag(calcDBH, n=1, default=NA, order_by=Tag),
         cii_min1=dplyr::lag(Cii, n=1, default=NA, order_by=Tag),
         tdif=difftime(Date2, lag(Date2, n=1, default=NA, order_by=Tag), units = "days"),
         inc=calcDBH-calcDBH_min1,
         inc_annual=(inc/as.numeric(tdif))*365,
         inc.time=Cno-lag(Cno),
         #removing increment value if the previous measure was before the last year
         #inc=ifelse(inc.time==1, inc, NA),
         inc_annual=ifelse(inc.time==2, inc_annual, NA),
         treeID=paste0("HKK.", Tag))%>%
  filter(!is.na(inc))%>%
  filter(!is.na(inc_annual))%>%
  ungroup()%>%
  dplyr::mutate(
    diainc.scaled = scale(inc_annual),
    out.inc = ifelse(abs(diainc.scaled)>3, 1, 0))%>%
  filter(out.inc==0)

@@

dendro_2015<- dendro.data.noouts %>%
  filter(Cno %in% c(11, 13, 15))%>%
  group_by(Tag) %>% #grouping by treeID to calculate increment
  mutate(calcDBH_min1=dplyr::lag(calcDBH, n=1, default=NA, order_by=Tag),
         cii_min1=dplyr::lag(Cii, n=1, default=NA, order_by=Tag),
         tdif=difftime(Date2, lag(Date2, n=1, default=NA, order_by=Tag), units = "days"),
         inc=calcDBH-calcDBH_min1,
         inc_annual=(inc/as.numeric(tdif))*365,
         inc.time=Cno-lag(Cno),
         #removing increment value if the previous measure was before the last year
         #inc=ifelse(inc.time==1, inc, NA),
         inc_annual=ifelse(inc.time==2, inc_annual, NA),
         treeID=paste0("HKK.", Tag))%>%
  filter(!is.na(inc))%>%
  filter(!is.na(inc_annual))%>%
  ungroup()%>%
  dplyr::mutate(
    diainc.scaled = scale(inc_annual),
    out.inc = ifelse(abs(diainc.scaled)>3, 1, 0))%>%
  filter(out.inc==0)

@@

dendro_2019<- dendro.data.noouts %>%
  filter(Cno %in% c(19, 21, 23))%>%
  group_by(Tag) %>% #grouping by treeID to calculate increment
  mutate(calcDBH_min1=dplyr::lag(calcDBH, n=1, default=NA, order_by=Tag),
         cii_min1=dplyr::lag(Cii, n=1, default=NA, order_by=Tag),
         tdif=difftime(Date2, lag(Date2, n=1, default=NA, order_by=Tag), units = "days"),
         inc=calcDBH-calcDBH_min1,
         inc_annual=(inc/as.numeric(tdif))*365,
         inc.time=Cno-lag(Cno),
         #removing increment value if the previous measure was before the last year
         #inc=ifelse(inc.time==1, inc, NA),
         inc_annual=ifelse(inc.time==2, inc_annual, NA),
         treeID=paste0("HKK.", Tag))%>%
  filter(!is.na(inc))%>%
  filter(!is.na(inc_annual))%>%
  ungroup()%>%
  dplyr::mutate(
    diainc.scaled = scale(inc_annual),
    out.inc = ifelse(abs(diainc.scaled)>3, 1, 0))%>%
  filter(out.inc==0)

quantiles_2018<-quantile(dendro_2019$inc_annual[dendro_2019$Cno==21], 
                         probs=c(0.25, 0.5, 0.75, 1))

quantiles_2019<-quantile(dendro_2019$inc_annual[dendro_2019$Cno==23], 
                         probs=c(0.25, 0.5, 0.75, 1))

dendro_2019_plot<-ggplot(data = dendro_2019, aes(x=inc_annual, group=factor(Cno), fill=factor(Cno), col=factor(Cno))) +
  geom_density(alpha=0.5) +
  scale_color_manual(values=c("forestgreen", "brown"))+
  scale_fill_manual(values=c("forestgreen", "brown"))+
  geom_vline(xintercept=quantiles_2018, col="forestgreen")+
  geom_text(aes(x=0.75, y=1, label="2018"), col="forestgreen", size=3)+
  geom_text(aes(x=0.75, y=1.5, label="2019"), col="brown", size=3)+
  geom_vline(xintercept=quantiles_2019, col="brown")+ 
  xlab("diameter increment (cm)") + ylab("density")+
  guides(col="none", fill="none")+
  theme_bw()

library(gridExtra)
png("../doc/display/increments_raw.png", width=8, height=4, units="in", res=300)
grid.arrange(dendro_2010_plot, dendro_2015_plot, dendro_2019_plot, nrow=1)
dev.off()

```{r long-term growth rates}

pat<-as.character(read.table("../data/HKK-dendro/git_pat.txt")[1,1])

remotes:::download("../data/HKK-dendro/stem1.RData",
                   "https://raw.githubusercontent.com/forestgeo/HKK-tree-growth-data/main/data/Census/raw_data/hkk.stem1.rdata",
                   auth_token = pat)
remotes:::download("../data/HKK-dendro/stem2.RData",
                   "https://raw.githubusercontent.com/forestgeo/HKK-tree-growth-data/main/data/Census/raw_data/hkk.stem2.rdata",
                   auth_token = pat)
remotes:::download("../data/HKK-dendro/stem3.RData",
                   "https://raw.githubusercontent.com/forestgeo/HKK-tree-growth-data/main/data/Census/raw_data/hkk.stem3.rdata",
                   auth_token = pat)
remotes:::download("../data/HKK-dendro/stem4.RData",
                   "https://raw.githubusercontent.com/forestgeo/HKK-tree-growth-data/main/data/Census/raw_data/hkk.stem4.rdata",
                   auth_token = pat)

load("../data/HKK-dendro/stem1.RData")
load("../data/HKK-dendro/stem2.RData")
load("../data/HKK-dendro/stem3.RData")
load("../data/HKK-dendro/stem4.RData")

hkk_census <- hkk_census %>%
  filter(status=="A")%>%
  filter(tag %in% dendro.data.noouts$Tag)%>%
  group_by(StemTag)%>%
  dplyr::mutate(dbhmin1 = dplyr::lag(dbh, n=1, default=NA, order_by=StemTag),
                inc_annual = (dbh-dbhmin1)/5)%>%
  filter(!is.na(dbhmin1))%>%
  dplyr::summarise(avg_inc = mean(inc_annual, na.rm=T),
                   sd_inc = sd(inc_annual, na.rm=T))

inc_census<-ggplot(data = hkk_census, aes(x=avg_inc)) +
  geom_density(alpha=0.5) +
  xlab("mean annual increment (cm)")+
  theme_bw()

png("../doc/display/census_increment.png", width=4, height=4, units="in", res=300)
inc_census
dev.off()

```{r drought resistance}

@@

#2010 drought--------

resistance_2010 <- dendro_2010 %>%
  group_by(Tag) %>% #grouping by treeID to calculate increment
  mutate(inc_min1=dplyr::lag(inc_annual, n=1, default=NA, order_by=Tag),
         resistance.div = inc_annual/inc_min1,
         resistance.dif = inc_annual-inc_min1,
         resistance.prop = (inc_annual-inc_min1)/inc_min1)%>%
  filter(!is.na(inc_min1), !is.infinite(resistance.div))

resistance_2010<-merge(resistance_2010, select(dec_williams, -"X"), by="sp", all.x=T)

@@

#2015 drought---------

resistance_2015 <- dendro_2015 %>%
  group_by(Tag) %>% #grouping by treeID to calculate increment
  mutate(inc_min1=dplyr::lag(inc_annual, n=1, default=NA, order_by=Tag),
         resistance.div = inc_annual/inc_min1,
         resistance.dif = inc_annual-inc_min1,
         resistance.prop = (inc_annual-inc_min1)/inc_min1)%>%
  filter(!is.na(inc_min1), !is.infinite(resistance.div))

resistance_2015<-merge(resistance_2015, select(dec_williams, -"X"), by="sp", all.x=T)

@@

#2019 drought--------------

resistance_2019 <- dendro_2019 %>%
  group_by(Tag) %>% #grouping by treeID to calculate increment
  mutate(inc_min1=dplyr::lag(inc_annual, n=1, default=NA, order_by=Tag),
         resistance.div = inc_annual/inc_min1,
         resistance.dif = inc_annual-inc_min1,
         resistance.prop = (inc_annual-inc_min1)/inc_min1)%>%
  filter(!is.na(inc_min1), !is.infinite(resistance.div))

resistance_2019<-merge(resistance_2019, select(dec_williams, -"X"), by="sp", all.x=T)

resistance_2019_div<-ggplot(data = resistance_2019, aes(x=resistance.div)) +
  geom_density(alpha=0.5) +
  theme_bw()

resistance_2019_dif<-ggplot(data = resistance_2019, aes(x=resistance.dif)) +
  geom_density(alpha=0.5) +
  theme_bw()

png("../doc/display/resistance_dif.png", width=4, height=8, units="in", res=300)
grid.arrange(resistance_2010_dif, resistance_2015_dif, resistance_2019_dif, nrow=3)
dev.off()

```{r resistance map}

quantiles_2010_dif<-quantile(resistance_2010$resistance.dif, 
                         probs=c(0.25, 0.5, 0.75, 1))

quantiles_2015_dif<-quantile(resistance_2015$resistance.dif, 
                         probs=c(0.25, 0.5, 0.75, 1))

quantiles_2019_dif<-quantile(resistance_2019$resistance.dif, 
                         probs=c(0.25, 0.5, 0.75, 1))

resistance_2010$quantile<-ifelse(resistance_2010$resistance.dif<quantiles_2010_dif[1], 1, ifelse(resistance_2010$resistance.dif<quantiles_2010_dif[2], 2,
            ifelse(resistance_2010$resistance.dif<quantiles_2010_dif[3], 3, 4))

                                 )

resistance_2015$quantile<-ifelse(resistance_2015$resistance.dif<quantiles_2015_dif[1], 1, ifelse(resistance_2015$resistance.dif<quantiles_2015_dif[2], 2,
            ifelse(resistance_2015$resistance.dif<quantiles_2015_dif[3], 3, 4))

                                 )

resistance_2019$quantile<-ifelse(resistance_2019$resistance.dif<quantiles_2019_dif[1], 1, ifelse(resistance_2019$resistance.dif<quantiles_2019_dif[2], 2,
            ifelse(resistance_2019$resistance.dif<quantiles_2019_dif[3], 3, 4))

                                 )

resistance_all<-rbind(resistance_2010, resistance_2015, resistance_2019)
resistance_all$yr<-c(rep(2010, nrow(resistance_2010)), rep(2015, nrow(resistance_2015)), rep(2019, nrow(resistance_2019)))

resistance_all<-merge(resistance_all, select(trees, -"SPCODE.UPDATE"), by="Tag")

resistance_map<-ggplot(resistance_all, aes(x=X, y=Y, size=as.numeric(DBH.04), col=quantile))+
  geom_point(alpha=0.5)+
   scale_size(range = c(.1, 10), name="DBH (cm)")+
  guides(color=guide_legend(title="resistance"))+
  scale_colour_gradient(low="red", high="blue") +
  facet_grid(yr~.)+
  theme_minimal()+
  coord_fixed()

png("../doc/display/resistance_map.png", width=8, height=10, units="in", res=300)
resistance_map
dev.off()
```
```{r basic model}

@@

model_2019<-glm(resistance.dif ~ calcDBH_min1 + cii_min1 + williams_dec+
                  calcDBH_min1*cii_min1 + 
                  calcDBH_min1*williams_dec+
                  cii_min1*williams_dec,
                data=resistance_2019)

summary(model_2019)

resistance_all<-rbind(resistance_2010, resistance_2015, resistance_2019)
resistance_all$yr<-c(rep(2010, nrow(resistance_2010)), rep(2015, nrow(resistance_2015)), rep(2019, nrow(resistance_2019)))

resistance_model<-resistance_all %>%
  select(resistance.dif, calcDBH_min1, cii_min1, williams_dec, yr, quantile)

@@

model_all_cat<-lme(quantile ~ calcDBH_min1 + cii_min1 + williams_dec+
                  calcDBH_min1*cii_min1 + 
                  calcDBH_min1*williams_dec+
                  cii_min1*williams_dec,
                random=~1|yr,
                data=resistance_model)

model_all_cat
anova(model_all_cat)
```

```{r predictions, echo=F}

library(sjPlot)

plot_model(model_all, type="pred", terms=c("calcDBH_min1","williams_dec"), show.data = T)
plot_model(model_all, type="pred", terms=c("cii_min1"), show.data = T)
plot_model(model_all, type="pred", terms=c("williams_dec"), show.data = T)
plot_model(model_all, type="pred", terms=c("calcDBH_min1"), show.data = T)

plot_model(model_all, type = "est")+
  theme_bw()+ylim(c(-0.25, 0.25))+
  geom_hline(yintercept = 0)

plot_model(model_all_cat, type = "est")+
  theme_bw()+ylim(c(-0.25, 0.25))+
  geom_hline(yintercept = 0)

plot_model(model_all_cat, type="pred", terms=c("calcDBH_min1","williams_dec"), show.data = T)

@@

@@
